Formal Synthesis of Digital Controllers for Observable Continuous Plants with Transient Performance Specifications

### 2018-01-23 TACAS2018/paper.tex
fixes on Cristinas comments

When our observer is implemented using a finite precision controller, new
problems arise due to the FWL effects.  Let ${\mat{A}_d}_{\langle C
\rangle}$, ${\mat{B}_d}_{\langle C \rangle}$ and ${\mat{C}_d}_{\langle C
\rangle}$ be the dynamics represented in $\mathbb{R}_{\langle C\rangle}^*$,
Equation \eqref{eq:observer_LTI_cf} becomes:
%
\begin{align}
\left [\begin{array}{c}\vec{x}'\\ \vec{x}_e' \end{array}\right]_{k+1}
=& \left [\begin{array}{cc}\mat{A}_{c}+\mat{B}_{e}&\mat{B}_{o}\\ \mat{A}_{e}&\mat{A}_{o}\end{array}\right]
\left [\begin{array}{c}\vec{x}'\\ \vec{x}_e' \end{array}\right]_k
+\left[\begin{array}{cc}\mat{T}_c{\mat{B}_d}_{\langle C \rangle}&\mat{0}\\\mat{0}&-\mat{T}_o\mat{L}_{\langle C \rangle}\end{array}\right]\left [\begin{array}{c} \vec{r}_k + \vec{\eta}_2\\ \vec{\eta}_1\end{array}\right]\nonumber\\
\vec{y}_k=&\mat{C}_d\mat{T}_c\vec{x}_k'% + \mat{D}_{d}\vec{u}_k
\label{eq:observer_LTI_FWL_cf}
\end{align}
where $\mat{T}_c$ and $\mat{T}_o$ trasfrom the system into canonical form~\cite{Astrom08} and
{\setlength{\abovedisplayskip}{2pt}
\setlength{\belowdisplayskip}{2pt}
\begin{align*}
\mat{A}_{c}=&\mat{T}_c\left(\mat{A}_d-\mat{B}_d\mat{K}\right)\mat{T}_c^{-1}&
\mat{A}_{o}=&\mat{T}_o\left({\mat{A}_d}_{\langle C \rangle}-\mat{L}_{\langle C \rangle}{\mat{C}_d}_{\langle C \rangle}\right)\mat{T}_o^{-1}\\
\mat{B}_{o}=&\mat{T}_c{\mat{B}_d}_{\langle C \rangle}\mat{K}_{\langle C \rangle}\mat{T}_o^{-1}&
\mat{B}_{e}=&\mat{T}_c(\mat{B}_d-{\mat{B}_d}_{\langle C \rangle})\mat{K}_{\langle C \rangle}\mat{T}_o^{-1}\\
\mat{A}_{e}=&\mat{T}_o\left(\mat{A}_d-{\mat{A}_d}_{\langle C \rangle}-\mat{L}_{\langle C \rangle}\mat{C}_e\right)\mat{T}_o^{-1}&
\mat{C}_{e}=&\mat{C}_d-{\mat{C}_d}_{\langle C \rangle}
\end{align*}
}
{\setlength{\abovedisplayskip}{0pt}
\begin{align*}
\mat{A}'=&\left [\begin{array}{cc}\mat{A}_{c}+\mat{B}_{e}&\mat{B}_{o}\\ \mat{A}_{e}&\mat{A}_{o}\end{array}\right]&
\mat{B}'=&\left[\begin{array}{cc}\mat{T}_c{\mat{B}_d}_{\langle C \rangle}&\mat{0}\\\mat{0}&-\mat{T}_o\mat{L}_{\langle C \rangle}\end{array}\right]
\end{align*}
}
The first effect of this digitalization is that $\mat{A}'$ is no longer upper
diagonal.  We must therefore compensate for the errors introduced in order
to find a stable solution.  We can use Gershgorin disks~\cite{van1996matrix} to limit the error
in the expected eigenvalues.  Let $\delta_o=\max(\norm{{\mat{A}_e}_{i,*}}_1)$
(\emph{i.e.} the largest row norm in $\mat{A}_e$) be the largest error bound
for any eigenvalue in $\mat{A}_e$.  Then $\forall i \in [1, ...\,,n],
|{\lambda_o}_{i}-\hat{\lambda_o}_{i}|<\delta_o$,where ${\lambda_o}_{i}$ are the
true eigenvalues of the observer and $\hat{\lambda_o}_{i}$ the eigenvalues of
the perturbed observer.  We note that in the case of stability we only care
about the error with respect to the unit circle, which is equivalent to
upper bounding the estimation of the largest eigenvalue.  Hence, given a
desired upper value $\overline{\lambda}$, we apply Jury's criterion on the
polynomial.
%
\begin{equation}
P_o'=\sum_{j=0}^p c_j\overline{\lambda'}^{p-j}s^j, \text{ where }  \overline{\lambda'}=(1+\delta_o)\overline{\lambda}
\end{equation}
%
In the case of the reachable eigenvalues, the procedure is the same except
that we obtain our bound $\delta_c=\max(\norm{{\mat{A}_e}_{*,i}+{\mat{B}_e}_{*,i}}_1)$
(\emph{i.e.} the largest column norm in $ \left[\begin{array}{c}\mat{A}_e\\ \mat{B}_e\end{array}\right]$) and
apply Jury's Criterion over
%
\begin{equation}
P_r'=\sum_{j=0}^p c_j\overline{\lambda'}^{p-j}s^j, \text{ where }  \overline{\lambda'}=(1+\delta_c)\overline{\lambda}
\end{equation}

@@

%-------------------------------
\subsection{Transient response} 
\label{ssec:transientspecification}
%-------------------------------
In many cases the speed of the response, i.e. how fast it converges to the set point, is a requirement of great importance. 
This is given as a settling time for the system to become stable.

@@

We build upon this paradigm to create CEGORS (Counter Example Guided, Optimization
Refined Synthesis) \note{not sure whether we need an acronym for this, we could just call it
CEGIS with optimization and abstraction refinement}.
Our approach is illustrated in Figure~\ref{fig:CEGORS} and consists of three main
phases:
  \note{explain broadly what each phase does}
\begin{itemize}
\item {\sc Abstract}: constructs of an unsound abstract model that over-approximates
the dynamics of the system.
\item {\sc synthesize}: generates a candidate controller and observer \note{are these obeying the stability spec?}.
\item {\sc model}
\item {\sc verify}: verifies the behaviour of the candidate controller and observer
  against a sound over-approximation of the concrete dynamics of the system for an infinite space over an infinite time horizon.
\end{itemize}

The verifier's abstraction uses a CEGAR loop in order to minimise 
over-approximation errors and provides counterexamples for both
discrete and continuous dynamics to feed a CEGIS loop on the synthesiser
waiting for a new candidate solution. \note{not sure about this sentence.
  Which is the CEGAR loop?}

The main advantage of this approach is two-fold:
\begin{itemize}
\item[1.] Optimizing the candidate solution in each
  iteration of the CEGIS loop by computing the local optimum
  enables us to obtain high quality counterexamples in the {\sc verify} phase.
  In turn, this has the potential to reduce the number of iteration of the CEGIS
  loop required to obtain a solution.
  \note{probably explain high quality counterexample}
\item[2.] We start with simpler (i.e. imprecise)
  abstractions that get
  refined into more complex ones as the procedure progresses. Thus, solutions
  that don't require a high level of precision are obtained fast. 
\end{itemize}

@@

During the optimization, the minimal abstraction used by the synthesiser almost
always ensures stability, so our objective function focuses on safety,
although we manage to do optimizations that simultaneously target
safety, stability and response time.
\note{Are all the
  candidates computed by {\sc synthesize} stable? State clearly if
  so. You say ``almost always''.}
This is done by making the
dynamics as convergent as possible while maintaining the verification
properties that exclude excesively convergent ones.

@@

The CEGORS loop for this approach is described next.
\paragraph{1. {\sc abstract}} creates an abstraction from the initial specificationas follows:
\begin{itemize}
\item Select a sample time that minimizes the magnitudes of the matrices used in both the
implementation and the abstraction. This ensures that a small bit vector size is required for
the synthesis process.
\item Build the observer matrices ${\mat{A}_d}_{\langle C \rangle}$, ${\mat{B}_d}_{\langle C \rangle}$
and ${\mat{C}_d}_{\langle C \rangle}$ using the truncation described in Section~\ref{sec:numeric_rep}.
\item Transform the system into canonical form and add nondeterminism for the errors
using the techniques described in Section~\ref{sec:cof_fwl_stability}. The transformation
matrices will be needed to obtain an FWL compliant controller and observer.
\item Compute the characteristic polynomial of the matrix $(\mat{A}_d-\mat{B}_d\mat{K}_{\langle C \rangle})$
as $P_a(z) = z^p+\sum_{i=1}^p{(a_i-k_i)z^{p-i}}$. 
\item Compute the characteristic polynomial of the matrix $({\mat{A}_d}_{\langle C \rangle}-\mat{L}_{\langle C \rangle}{\mat{C}_d}_{\langle C \rangle})$
as $P_l(z) = z^p+\sum_{i=1}^p{(l_i-a_i)z^{p-i}}$. 
\item Create a stability specification based on the characteristic polynomials, perturbation
error ($\mat{A}_e$,$\mat{B}_e$), and response time $\phi_{stability}(P_a,P_l,\overline{\lambda})$
as described in Section~\ref{sec:transientspecification} \note{broken reference}.
\item Calculate the noise sets $\eta_1$ and $\eta_2$ from the equations
described in Section~\ref{sec:FWL_effects} \note{broken reference}.
\item Calculate a set of initial bounds on $\mat{K}$, $\phi_\mathit{init}^{K}$,
based on the input constraints.
$(\phi_\mathit{init} \wedge \phi_\mathit{input} \wedge u_k=-\mat{K} \vec{x}_k)
\Rightarrow \phi_\mathit{init}^{K}$.
These bounds will be used by the {\sc synthesize} phase to reduce the size of the solution space. 
\item Select an initial width for the bit vectors representing the controller and the plant $\phi_{\langle C \rangle}, \phi_{\langle P \rangle}$. We will increase either precision as needed.
\end{itemize}
\note{This paragraph needs to be moved before the itemize and it needs to be improved: give an intuition for what $\phi_{static}$ and $\phi_{synth}$ represent, e.g. the system's abstraction + stability spec + safety spec...}
There are two abstractions created by these specifications: $\phi_{synth}$ which is given to the synthesiser and $\phi_{static}$ which is given to the synthesis refinement loop. These abstractions will be refined independently, so we will not detail what each will contain at a given stage. We only note that at all times $\phi_{static} \models \phi_{synth}$.
%
%\item
\paragraph{2. {\sc synthesize}} generates a candidate controller $\mat{K}$
and observer $\mat{L}$ that satisfy
$\phi_{synth}$ by invoking a SAT solver.
There are two exit conditions:
If there is no candidate solution we return FAIL and signal {\sc Abstract}. Based on the current abstraction,
{\sc  Abstract} may choose to increase the precision, refine the abstraction, or return a synthesis failure.
If we have a candidate solution we will feed it to the verifier chain in order to validate it.
%
%\item
\paragraph{3. {\sc Verify Synthesis}} verifies $\phi_{static}$ of  both the
controller and the observer.  If one of these results does not satisfy this tighter
specification, it feeds back to the {\sc Optimise} phase.
If the candidate is valid, we feed it to the next {\sc verify} stage, {\sc Discrete Time}.

%\item
\paragraph{4. {\sc Optimise}} computes the local optimum in the neighborhood of
the current candidate solution. \note{is this really a local optimum?}
% starts by taking $\mat{K},\mat{L} \not \models \phi_{static}$
% and synthesises $\mat{K}_{\langle C \rangle} = \mat{K}+\mat{\delta}_K, \mat{L}_{\langle C \rangle}=\mat{L}+\mat{\delta}_L$,
% where $\mat{K}_{\langle C \rangle}, \mat{L}_{\langle C \rangle} \models \phi_{static}$.  
% If $\mat{K},\mat{L} \models \phi_{static}$ but we have a counterexample for $\mat{K},\mat{L} \models \phi$,
% then we proceed to optimise the controller.
%
Our optimization is iterative, following two stages.  The first stage seeks to
make the dynamics as convergent as possible, since convergent dynamics are
less likely to stray outside of the safe region (we say less likely because the
change in the eigenvectors changes the shape of the space and may cause some 
trajectories to go further).
\note{Can we formally define the objective function?}

In this case, we take the largest eigenvalue and reduce it by a factor
(smaller than .5) that still satisfies $\phi_{static}$.  The resulting
observer controller is then verified against the full specification
\note{full specification?}
and the maximum deviation from the safe region is measured (i.e. maximum
distance from the discovered reach set to the safety region, measured
perpendicularly to any of the faces of the safe region).  While this
number is monotonically decreasing we continue to iterate until it
reaches a negative value (i.e. safe). If the number increases, we
change our objective function to the second stage.
\note{again, formal description? what is the objective function in this stage?}

In the second stage, at each iteration we take the complex eigenvalue
with the largest phase (largest angle in the complex space), and reduce its
imaginary component while maintaining the specification $\phi_{static}$.
The process ends when no imaginary part can be further reduced.
The resulting closed loop form converges faster than the original and may have a
smaller rotation, which facilitates meeting the safety specification
\note{$\phi_{safety}$?}
(we note
that real convergent eigenvalues are trivially safe).

@@

%-------------------------------
\subsection{Transient response} 
\label{sec:transientspecification}
%-------------------------------
\begin{theorem}
Given a set of eigenvalues for a discrete time model of an $p^{th}$ order system $\lambda_i =e^{\omega_i}, i \in [1, ...\,,p]$, the settling time of the system is upper-bounded by
\begin{equation}
t_s<-log_{\overline{\lambda}}({P_s}), \text{ where }  \overline{\lambda} = \max(|\lambda_i|) \text{ and } \overline{\lambda}<1,\ \forall i \in [1, ...\,,p]
\label{eq:set_time}
\end{equation}
where $0\leq P_s \leq 1$ is the desired percentage of the target within which the output must remain after $t_s$.
\end{theorem}
\begin{proof}
In a SISO system, there are two values that typically define the
response-time of the system (using a step response).
%
\begin{enumerate}
\item The \emph{rise time} is defined as the time it takes the output to reach a percentage (eg $90\%$) of the target value given a step chnage in the input.
\item The \emph{settling time} is defined as the time it takes the output to stabilize within a margin of the target value (eg $10\%$).
\end{enumerate} 
%
In the case of first and second order systems, the decay is governed by the equations 
%
\begin{align*}
&y_{decay}=
&\left\{
\begin{array}{cc}
e^{-|\omega| t}& 1^{st} order\\
e^{\zeta |\omega| t}\left(\frac{\sin\left(|\omega| t\sqrt{1-\zeta^2}+\sin^{-1}\sqrt{1-\zeta^2}\right)}{\sqrt{1-\zeta^2}} \right)&\zeta<1\\
e^{-|\omega| t}-|\omega| t e^{-|\omega| t}&\zeta=1\\
\frac{|\omega| e^{-|\omega_2| t}-|\omega_2| e^{-|\omega| t}}{|\omega|-|\omega_2|}&\zeta>1\end{array}\right.
\end{align*}
where $|\omega|$ and $|\omega_2|$ are the magnitudes of the eigenvalues (roots) of the system and  $\zeta=-\cos(\theta)$ where $\theta$ is the polar angle of the conjugate pair if it exists ($\zeta \geq 1$ if it doesn't).
These equations become increasingly complicated for higher order systems. However, we note their strong dependency on the magnitude of the eigenvalues, and use this property to define an upperbound on the decay function.
All of the above equations will decay faster than $e^{-|\omega| t}$ given $|\omega| < |\omega_2|$, hence we may use this function as our upper bound.

>>>>

When our observer is implemented using a finite precision controller, new
problems arise due to the FWL effects.  Let ${\mat{A}_d}_{\langle C
\rangle}$, ${\mat{B}_d}_{\langle C \rangle}$ and ${\mat{C}_d}_{\langle C
\rangle}$ be the dynamics represented in $\mathbb{R}_{\langle C\rangle}^*$,
Equation \eqref{eq:observer_LTI_cf} becomes:
%
\begin{align}
\left [\begin{array}{c}\vec{x}'\\ \vec{x}_e' \end{array}\right]_{k+1}
=& \left [\begin{array}{cc}\mat{A}_{c}+\mat{B}_{e}&\mat{B}_{o}\\ \mat{A}_{e}&\mat{A}_{o}\end{array}\right]
\left [\begin{array}{c}\vec{x}'\\ \vec{x}_e' \end{array}\right]_k
+\left[\begin{array}{cc}\mat{T}_c{\mat{B}_d}_{\langle C \rangle}&\mat{0}\\\mat{0}&-\mat{T}_o\mat{L}_{\langle C \rangle}\end{array}\right]\left [\begin{array}{c} \vec{r}_k + \vec{\eta}_2\\ \vec{\eta}_1\end{array}\right]\nonumber\\
\vec{y}_k=&\mat{C}_d\mat{T}_c\vec{x}_k'% + \mat{D}_{d}\vec{u}_k
\label{eq:observer_LTI_FWL_cf}
\end{align}
where $\mat{T}_c$ and $\mat{T}_o$ trasfrom the system into canonical form~\cite{Astrom08} and
{\setlength{\abovedisplayskip}{2pt}
\setlength{\belowdisplayskip}{2pt}
\begin{align*}
\mat{A}_{c}=&\mat{T}_c\left(\mat{A}_d-\mat{B}_d\mat{K}\right)\mat{T}_c^{-1}&
\mat{A}_{o}=&\mat{T}_o\left({\mat{A}_d}_{\langle C \rangle}-\mat{L}_{\langle C \rangle}{\mat{C}_d}_{\langle C \rangle}\right)\mat{T}_o^{-1}\\
\mat{B}_{o}=&\mat{T}_c{\mat{B}_d}_{\langle C \rangle}\mat{K}_{\langle C \rangle}\mat{T}_o^{-1}&
\mat{B}_{e}=&\mat{T}_c(\mat{B}_d-{\mat{B}_d}_{\langle C \rangle})\mat{K}_{\langle C \rangle}\mat{T}_o^{-1}\\
\mat{A}_{e}=&\mat{T}_o\left(\mat{A}_d-{\mat{A}_d}_{\langle C \rangle}-\mat{L}_{\langle C \rangle}\mat{C}_e\right)\mat{T}_o^{-1}&
\mat{C}_{e}=&\mat{C}_d-{\mat{C}_d}_{\langle C \rangle}
\end{align*}
}
{\setlength{\abovedisplayskip}{0pt}
\begin{align*}
\mat{A}'=&\left [\begin{array}{cc}\mat{A}_{c}+\mat{B}_{e}&\mat{B}_{o}\\ \mat{A}_{e}&\mat{A}_{o}\end{array}\right]&
\mat{B}'=&\left[\begin{array}{cc}\mat{T}_c{\mat{B}_d}_{\langle C \rangle}&\mat{0}\\\mat{0}&-\mat{T}_o\mat{L}_{\langle C \rangle}\end{array}\right]\\
\eta_1=&\nu_1+\nu_1'+\nu_t& \eta_2=&\nu_2+\nu_2'
\end{align*}
}
where $\nu_1$ and $\nu_2$ are the quantization noises introduced by the ADC and the DAC, $\nu_t$ is the noise introduced by the delay, and $\nu_1'$ and $\nu_2'$ are quantization noises caused by roundoffs in the FWL operations (see Section~\ref{sec:quantization-noise}).

The first effect of this digitalization is that $\mat{A}'$ is no longer upper
diagonal.  We must therefore compensate for the errors introduced in order
to find a stable solution.  We can use Gershgorin disks~\cite{van1996matrix} to limit the error
in the expected eigenvalues.  Let $\delta_o=\max(\norm{{\mat{A}_e}_{i,*}}_1)$
(\emph{i.e.} the largest row norm in $\mat{A}_e$) be the largest error bound
for any eigenvalue in $\mat{A}_e$.  Then $\forall i \in [1, ...\,,n],
|{\lambda_o}_{i}-\hat{\lambda_o}_{i}|<\delta_o$,where ${\lambda_o}_{i}$ are the
true eigenvalues of the observer and $\hat{\lambda_o}_{i}$ the eigenvalues of
the perturbed observer.  We note that in the case of stability we only care
about the error with respect to the unit circle, which is equivalent to
upper bounding the estimation of the largest eigenvalue.  Hence, given a
desired upper value $\overline{\lambda}$, we apply Jury's criterion on the
polynomial.
%
\begin{equation}
P_o'=\sum_{j=0}^p c_j\overline{\lambda'}^{p-j}s^j, \text{ where }  \overline{\lambda'}=(1+\delta_o)\overline{\lambda}
\end{equation}
%
In the case of the reachable eigenvalues, the procedure is the same except
that we obtain our bound $\delta_c=\max(\norm{{\mat{A}_e}_{*,i}+{\mat{B}_e}_{*,i}}_1)$
(\emph{i.e.} the largest column norm in $ \left[\begin{array}{c}\mat{A}_e\\ \mat{B}_e\end{array}\right]$) and
apply Jury's Criterion over
%
\begin{equation}
P_r'=\sum_{j=0}^p c_j\overline{\lambda'}^{p-j}s^j, \text{ where }  \overline{\lambda'}=(1+\delta_c)\overline{\lambda}
\end{equation}

@@

%-------------------------------
\subsection{Transient response} 
\label{sec:transientspecification}
%-------------------------------
In many cases the speed of the response, i.e. how fast it converges to the set point, is a requirement of great importance. 
This is given as a settling time for the system to become stable.

@@

We build upon this paradigm to create CEGORS (Counter Example Guided, Optimization
Refined Synthesis) \note{not sure whether we need an acronym for this, we could just call it
CEGIS with optimization and abstraction refinement}.
Our approach is illustrated in Figure~\ref{fig:CEGORS} and consists of three main
phases:
  \note{explain broadly what each phase does}
\begin{itemize}
\item {\sc Abstract}: constructs of an unsound abstract model that over-approximates
the dynamics of the system.
\item {\sc synthesize}: generates a candidate controller and observer.  Since
  we are using coarse approximations, these do not necessarily meet any portion
  of the specification, but either through optimization or refinement become
  more precise (therefore more likely to meet the specification) with each iteration.
\item {\sc model}: refines or redefines the abstractions needed by {\sc Abstract}
  in order to reduce the probability of failure in the next iteration.
\item {\sc verify}: verifies the behaviour of the candidate controller and observer
  against a sound over-approximation of the concrete dynamics of the system for an infinite space over an infinite time horizon.
\end{itemize}

%The verifier's abstraction uses a CEGAR loop in order to minimise 
%over-approximation errors and provides counterexamples for both
%discrete and continuous dynamics to feed a CEGIS loop on the synthesiser
%waiting for a new candidate solution. \note{not sure about this sentence.
%  Which is the CEGAR loop?}

The main advantage of this approach is two-fold:
\begin{itemize}
\item[1.] Optimizing the candidate solution in each iteration of the CEGIS loop
  enables us to obtain better candidates that are more likely to meet the
  specification or fail by a smaller margin.  This has the potential to reduce
  the number of iteration of the CEGIS loop required to obtain a solution.
\item[2.] We start with simpler (i.e. imprecise) abstractions that get
  refined into more complex ones as the procedure progresses. Thus, solutions
  that don't require a high level of precision are obtained fast. 
\end{itemize}

@@

We use two objective functions for optimization.  Both aim at reducing the
size of the reach space of the closed loop system while maintaining other
verification properties.  The first one aims at making the system as
convergent as possible.  This does not only ensure stability, but reduces
the response time of the system, moving away from the unsafe states faster.
The second one seeks to minimize rotations.  If the system rotates slower, it
is less likely to reach unsafe states before converging.
In both cases, the shape of the eigenspace may affect us negativey, so we
use a measure of the size of the reach space to guide our optimization.

@@

The CEGORS loop for this approach is described next.
\paragraph{1. {\sc abstract}} creates an abstraction from the initial specificationas follows:
\begin{itemize}
\item Select a sample time that minimizes the magnitudes of the matrices used in both the
implementation and the abstraction. This ensures that a small bit vector size is required for
the synthesis process.
\item Build the observer matrices ${\mat{A}_d}_{\langle C \rangle}$, ${\mat{B}_d}_{\langle C \rangle}$
and ${\mat{C}_d}_{\langle C \rangle}$ using the truncation described in Section~\ref{sec:numeric_rep}.
\item Transform the system into canonical form and add nondeterminism for the errors
using the techniques described in Section~\ref{sec:cof_fwl_stability}. The transformation
matrices will be needed to obtain an FWL compliant controller and observer.
\item Compute the characteristic polynomial of the matrix $(\mat{A}_d-\mat{B}_d\mat{K}_{\langle C \rangle})$
as $P_a(z) = z^p+\sum_{i=1}^p{(a_i-k_i)z^{p-i}}$. 
\item Compute the characteristic polynomial of the matrix $({\mat{A}_d}_{\langle C \rangle}-\mat{L}_{\langle C \rangle}{\mat{C}_d}_{\langle C \rangle})$
as $P_l(z) = z^p+\sum_{i=1}^p{(l_i-a_i)z^{p-i}}$. 
\item Create a stability specification based on the characteristic polynomials, perturbation
error ($\mat{A}_e$,$\mat{B}_e$), and response time $\phi_{stability}(P_a,P_l,\overline{\lambda})$
as described in Section~\ref{sec:transientspecification}.
\item Calculate the noise sets $\eta_1$ and $\eta_2$ from the equations
described in Section~\ref{sec:observable}.
\item Calculate a set of initial bounds on $\mat{K}$, $\phi_\mathit{init}^{K}$,
based on the input constraints.
$(\phi_\mathit{init} \wedge \phi_\mathit{input} \wedge u_k=-\mat{K} \vec{x}_k)
\Rightarrow \phi_\mathit{init}^{K}$.
These bounds will be used by the {\sc synthesize} phase to reduce the size of the solution space. 
\item Select an initial width for the bit vectors representing the controller and the plant $\phi_{\langle C \rangle}, \phi_{\langle P \rangle}$. We will increase either precision as needed.
\end{itemize}
\note{This paragraph needs to be moved before the itemize and it needs to be improved: give an intuition for what $\phi_{static}$ and $\phi_{synth}$ represent, e.g. the system's abstraction + stability spec + safety spec...}
There are two abstractions created by these specifications: $\phi_{synth}$ which is given to the synthesiser and $\phi_{static}$ which is given to the synthesis refinement loop. These abstractions will be refined independently, so we will not detail what each will contain at a given stage. We only note that at all times $\phi_{static} \models \phi_{synth}$.
%
%\item
\paragraph{2. {\sc synthesize}} generates a candidate controller $\mat{K}$
and observer $\mat{L}$ that satisfy
$\phi_{synth}$ by invoking a SAT solver.
There are two exit conditions:
If there is no candidate solution we return FAIL and signal {\sc Abstract}. Based on the current abstraction,
{\sc  Abstract} may choose to increase the precision, refine the abstraction, or return a synthesis failure.
If we have a candidate solution we will feed it to the verifier chain in order to validate it.
%
%\item
\paragraph{3. {\sc Verify Synthesis}} verifies $\phi_{static}$ of  both the
controller and the observer.  If one of these results does not satisfy this tighter
specification, it feeds back to the {\sc Optimise} phase.
If the candidate is valid, we feed it to the next {\sc verify} stage, {\sc Discrete Time}.

%\item
\paragraph{4. {\sc Optimise}} refines the candidate solution towards
a local optimum.  We seek to minimize the distance of the farthest point in the
reach space to the closest point to it in the safe region
\begin{equation}
e=max(\{min(\{\vec{x}-\vec{g} \mid \vec{g}\in G\}) \mid \vec{x} \in \hat{X}_\infty \}.
\label{eq:objective}
\end{equation}

Our optimization is iterative, following two stages.  The first stage seeks to
make the dynamics as convergent as possible, since convergent dynamics are
less likely to stray outside of the safe region (we say less likely because the
change in the eigenvectors changes the shape of the space and may cause some 
trajectories to go further).

In this case, we take the largest eigenvalue and reduce it by a factor $f$
that still satisfies the controller constraints ($f \overline{\lambda} \models \phi_{static}$).
The resulting observer controller is then verified against the safety specification $\phi_{safety}$
and the objective $e$ is measured (Equation \eqref{eq:objective}).
If for each iteration $k$, the objective $e_k<e_{k-1}$, we continue to iterate until 
$e_k \leq 0$ (i.e. we found a safe controller), otherwise we move to the second stage.

In the second stage, we take the complex eigenvalue
with the largest angle ($max\{|tan^{-1}\left(\frac{im(\lambda)}{re(\lambda)}\right)| \mid \lambda \in \Lambda\}$)
and reduce its imaginary component only  ($re(\lambda) + f im(\lambda) \models \phi_{static}$).
The process ends when no imaginary part can be further reduced.
The resulting closed loop form converges faster than the original and has a
smaller rotation, which facilitates meeting the safety specification $\phi_{safety}$
(we note that real convergent eigenvalues are trivially safe).

@@

%-------------------------------
\subsection{Transient response} 
\label{sec:transientspecificationproof}
%-------------------------------
\begin{theorem}
Given a set of eigenvalues for a discrete time model of an $p^{th}$ order system $\lambda_i =e^{\omega_i}, i \in [1, ...\,,p]$, the settling time of the system is upper-bounded by
\begin{equation}
t_s<-log_{\overline{\lambda}}({P_s}), \text{ where }  \overline{\lambda} = \max(|\lambda_i|) \text{ and } \overline{\lambda}<1,\ \forall i \in [1, ...\,,p]
\label{eq:set_time}
\end{equation}
where $0\leq P_s \leq 1$ is the desired percentage of the target within which the output must remain after $t_s$.
\end{theorem}
\begin{proof}
In a SISO system, there are two values that typically define the
response-time of the system (using a step response).
%
\begin{enumerate}
\item The \emph{rise time} is defined as the time it takes the output to reach a percentage (eg $90\%$) of the target value given a step chnage in the input.
\item The \emph{settling time} is defined as the time it takes the output to stabilize within a margin of the target value (eg $10\%$).
\end{enumerate} 
%
In the case of first and second order systems, the decay is governed by the equations 
%
\begin{align*}
&y_{decay}=
&\left\{
\begin{array}{cc}
e^{-|\omega| t}& 1^{st} order\\
e^{\zeta |\omega| t}\left(\frac{\sin\left(|\omega| t\sqrt{1-\zeta^2}+\sin^{-1}\sqrt{1-\zeta^2}\right)}{\sqrt{1-\zeta^2}} \right)&\zeta<1\\
e^{-|\omega| t}-|\omega| t e^{-|\omega| t}&\zeta=1\\
\frac{|\omega| e^{-|\omega_2| t}-|\omega_2| e^{-|\omega| t}}{|\omega|-|\omega_2|}&\zeta>1\end{array}\right.
\end{align*}
where $|\omega|$ and $|\omega_2|$ are the magnitudes of the eigenvalues (roots) of the system and  $\zeta=-\cos(\theta)$ where $\theta$ is the polar angle of the conjugate pair if it exists ($\zeta \geq 1$ if it doesn't).
These equations become increasingly complicated for higher order systems. However, we note their strong dependency on the magnitude of the eigenvalues, and use this property to define an upperbound on the decay function.
All of the above equations will decay faster than $e^{-|\omega| t}$ given $|\omega| < |\omega_2|$, hence we may use this function as our upper bound.
